---
title: "R-Ladies Melbourne highlights from 2018!"
author: "Anna Quaglieri, 2018 Main organiser"
date: "January 2018"
output:
  html_document:
    includes:
      after_body:
      - header.html
      - footer.html
    theme: cosmo
    toc: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(magick)
library(png)
library(grid)
library(emo)
#library(icon)
```

# Where are we at?

The R-Ladies Melbourne chapter was founded in September 2016 by [Göknur Giner](https://twitter.com/goknurginer) and [Sepideh Foroutan](https://twitter.com/S_Foroutan). Together with [Roxane Legaie](https://twitter.com/RoxaneLegaie), [Saskia Freytag](https://twitter.com/trashystats), Alex Garnham, [Soroor Zadeh](https://twitter.com/SoroorHediyeh) and [myself](https://twitter.com/annaquagli) they organized the first committee meeting at the Walter and Eliza Hall Institute and the [launch event](https://www.meetup.com/R-Ladies-Melbourne/events/234683544/) took of in October! 

Time flies, and it’s already been two years since our launch event. And just as a holiday ends and recollections of all the best moments begin to take form in your mind (can’t really think of any bad one!), I decided to take some time to collect the pieces of a special year for the Aussie R and R-Ladies community.

We celebrated our first anniversary a little over a year ago, and reached nearly 600 members on Meetup. In 2018 we celebrated again on the 17th of October and reached over 1000 members! We continued to host monthly events from February to December, and couldn’t really be happier with the consistently strong participation at our events across the year. Of course, we can’t claim that we ordinarily would attract 500 members at a given event (we would really need to have hired a bigger space…) but the steady increase in member turnout over the year is a reminder to us that this group is still needed and continues to attract more interest!


```{r fig.cap="Fig1. New members on meetup and events from October 2017 to October 2018.", echo = FALSE,fig.asp=0.50}
img <- readPNG("AllPhotos/time_series.png")
grid.raster(img)
```

<br>
  
  ```{r fig.cap="Fig2. Speakers from November 2017 to October 2018", echo = FALSE,fig.asp=0.50}
img <- readPNG("AllPhotos/faces2018.png")
grid.raster(img)
```


You can find all the materials produced and used across these two past years at our GitHub repository https://github.com/R-LadiesMelbourne/Links-to-Events-Repos and find videos of seminars at our R-Ladies Melbourne YouTube channel.

<br>

# A year of FIRST time things
  
2018 marked a number of breakthroughs for our group.  For the first time we:
  
*	Participated in the useR! conference in Brisbane and granted scholarships to participants.
*	Organised a mini conference with lightning talks and a panel discussion to celebrate our two-year anniversary.
*	Organised a lunch meetup, which hopefully is the first of many to come!
*	Decided, studied and planned to become a not-for-profit organisation (happening February 2019!) (I will expand on this topic in a separate post!)

# useR! 2018

## Getting all together!

useR! 2018 was such an amazing experience for me personally and I believe for all of the R Australian community as well. I found myself in the most ideal place and it was a refreshing five days of immersion with people that love and work towards common goals. Also, I had previously only seen a lot of those people on Twitter and never met in person! It was cool to chat at the same table with R-Ladies from various parts of Australia and the whole world! We had the opportunity to discuss amongst ourselves about our experience in organising events, how everyone goes about looking for sponsors, and suggestions about sourcing speakers for events.

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Here are amazing <a href="https://twitter.com/hashtag/RLadies?src=hash&amp;ref_src=twsrc%5Etfw">#RLadies</a> from around the world having dinner after an excellent day at <a href="https://twitter.com/useR2018_conf?ref_src=twsrc%5Etfw">@useR2018_conf</a> <a href="https://twitter.com/RLadiesGlobal?ref_src=twsrc%5Etfw">@RLadiesGlobal</a> <a href="https://twitter.com/RLadiesBrisbane?ref_src=twsrc%5Etfw">@RLadiesBrisbane</a> <a href="https://twitter.com/RLadiesSydney?ref_src=twsrc%5Etfw">@RLadiesSydney</a> <a href="https://twitter.com/RLadiesIstanbul?ref_src=twsrc%5Etfw">@RLadiesIstanbul</a> <a href="https://twitter.com/RLadiesIzmir?ref_src=twsrc%5Etfw">@RLadiesIzmir</a> <a href="https://twitter.com/RLadiesAKL?ref_src=twsrc%5Etfw">@RLadiesAKL</a> <a href="https://twitter.com/RLadiesDC?ref_src=twsrc%5Etfw">@RLadiesDC</a> <a href="https://twitter.com/RLadiesRemote?ref_src=twsrc%5Etfw">@RLadiesRemote</a> <a href="https://twitter.com/RLadiesMVD?ref_src=twsrc%5Etfw">@RLadiesMVD</a> <a href="https://twitter.com/hashtag/rladies?src=hash&amp;ref_src=twsrc%5Etfw">#rladies</a> <a href="https://twitter.com/hashtag/useR2018?src=hash&amp;ref_src=twsrc%5Etfw">#useR2018</a> <a href="https://twitter.com/hashtag/rstat?src=hash&amp;ref_src=twsrc%5Etfw">#rstat</a> <a href="https://t.co/aZoSuAU0Gi">pic.twitter.com/aZoSuAU0Gi</a></p>&mdash; R-Ladies Melbourne (@RLadiesMelb) <a href="https://twitter.com/RLadiesMelb/status/1016604921154584576?ref_src=twsrc%5Etfw">July 10, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
  
We also managed to organise a lunch for R-Ladies and conference newbies to come together and give support to newcomers and to people that came alone to the conference. A super R-Ladies cake was also there for the occasion!
  
  
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Thanks all of the amazing women coming around to R-ladies <a href="https://twitter.com/hashtag/useR2018?src=hash&amp;ref_src=twsrc%5Etfw">#useR2018</a> lunch today! <a href="https://t.co/7cwJUZi8mg">pic.twitter.com/7cwJUZi8mg</a></p>&mdash; R-Ladies Melbourne (@RLadiesMelb) <a href="https://twitter.com/RLadiesMelb/status/1016883039241043973?ref_src=twsrc%5Etfw">July 11, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
  
  
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">R-ladies lunch at room P11! Come and join us! <a href="https://twitter.com/hashtag/useR2018?src=hash&amp;ref_src=twsrc%5Etfw">#useR2018</a> <a href="https://t.co/bqvGb7CmPv">pic.twitter.com/bqvGb7CmPv</a></p>&mdash; R-Ladies Melbourne (@RLadiesMelb) <a href="https://twitter.com/RLadiesMelb/status/1016875124610351104?ref_src=twsrc%5Etfw">July 11, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
  

## We love our R babies
  
For the first time the conference was held in our hemisphere and this gave us the opportunity to actively contribute under [Di Cook](https://research.monash.edu/en/persons/dianne-cook)’s direction.  Some of us helped with some parts of the organization of the conference like reviewing abstracts and scholarship applications or organising the childcare for the conference. Göknur Giner helped with the childcare organisation and as a result she was able, along with the other parents, to attend the conference with her baby and letting her go off to childcare to listen to talks and tutorials when needed. This is an opportunity that every conference should offer! Imagine how many amazing parents who keep missing out if we don’t?
  
  
<blockquote class="twitter-tweet" data-lang="en"><p lang="und" dir="ltr"><a href="https://twitter.com/hashtag/hexwall?src=hash&amp;ref_src=twsrc%5Etfw">#hexwall</a> <a href="https://twitter.com/hashtag/useR2018?src=hash&amp;ref_src=twsrc%5Etfw">#useR2018</a> <a href="https://twitter.com/hashtag/maya?src=hash&amp;ref_src=twsrc%5Etfw">#maya</a> <a href="https://twitter.com/hashtag/childfriendlyconferences?src=hash&amp;ref_src=twsrc%5Etfw">#childfriendlyconferences</a> <a href="https://t.co/JN55krtZe5">pic.twitter.com/JN55krtZe5</a></p>&mdash; Göknur Giner (@goknurginer) <a href="https://twitter.com/goknurginer/status/1017362105949802497?ref_src=twsrc%5Etfw">July 12, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
  
## CSL gives us more than money!
  
As R-Ladies we had the chance to sponsor the conference providing two $500 scholarships which allowed two R-Ladies living in Melbourne, Dewi Angraini and Lucy Liu, to take part in the conference. Increasing gender and diversity balance at R conferences is, after all, one of the core aims of the whole R-Ladies movement. Having the chance to actively contribute to it made us feel super powerful and hopeful! We are enourmously thankful to our sponsor CSL from Melbourne which is giving us this and many other opportunities to grow and pursue the change in which we believe. Below are the two enthusiastic R-Ladies who were granted the scholarships. They also shared with us below a few words about what that travel grant meant for them!
  
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Congratulations again to the two winners of our scholarship for <a href="https://twitter.com/hashtag/useR2018?src=hash&amp;ref_src=twsrc%5Etfw">#useR2018</a> <a href="https://twitter.com/lucyleeow?ref_src=twsrc%5Etfw">@lucyleeow</a> and Dewi, finally here, with all the rest! <a href="https://t.co/3Q94qAaVjt">pic.twitter.com/3Q94qAaVjt</a></p>&mdash; R-Ladies Melbourne (@RLadiesMelb) <a href="https://twitter.com/RLadiesMelb/status/1016892581483954176?ref_src=twsrc%5Etfw">July 11, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
  
**Lucy's useR! 2018**

*"Would like to thank the RLadies Melbourne group for offering me a scholarship to attend the useR!2018 conference. The conference was an amazing showcase of the wide range of fields where R is used and the diversity of R users. The R community is known for being welcoming and diverse and the conference gave me a great opportunity to network these talented people. The interesting discussion on the variety of simple ways to help the community (e.g. via editing documentation, offering to write vignettes) has also inspired me to start contributing. I would not have been able to attend without the help of this scholarship and I am very grateful."* 

**Dewi's useR! 2018**
  
*"As a person who has background of Mathematics and Statistics, I thought I was too late to learn R and it was embarrassing that I just started learning R in the final year of my PhD. However, my great passion and intention to bring the knowledge of R back to my home country and teach my students about R has been significantly appreciated by the R-Ladies Melbourne by granting me the domestic travel scholarship to attend useR! 2018 in Brisbane. The conference gave me chance to engage with people from the R community from different countries, including  my home country Indonesia, and pushed me to further enlarge the R community in Indonesia which apparently remains limited, particularly in rural areas. I put a big hope that R-Ladies Melbourne can continuously support us in initiating the expansion of R community in Indonesia".*


# Two years anniversary
  
## Lightning talks and the best R prices!
  
To celebrate the special occasion we organised the first series of lightning talks where we asked our members to share with us their favourite R tips! They are now all collected and showcased together here https://rladies-melbourne-tips.netlify.com/ (strongly suggest to check it out!)! To this day I still read through it and have started using interesting features like the `DT::datatable` to display tables on HTML pages, allowing for filtering, sorting and easy searching!
  
This sounds all very exciting, but actually from day one up until almost the last few days we weren’t really sure that it was going to work. We didn’t have any submissions for lighning talks until the final three days. And this is not because people are lazy! When you organise events and you hope to attract volunteers to speak at evening events which are not official conferences and after a hard day at work… you need to realise that it can be a big thing to ask! However, we didn’t give up and that was the most important thing. We supported each other, we knew we might have had to share something ourselves as organisers and we just kept trying! When trying to organise something new, it’s important not to give up. Do everything that you can to keep your plan and your initial idea going. Most times, I believe, you will be rewarded. Indeed, in the end we were extremely happy to see enthusiatic young women come and share their expertise with all of us. We must admit that we also had amazing 3D printed R earrings to attract speakers!
  
  
<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">😮 Did you know you could win 🏆 these fab 💃 uniquely made eaRRings (purple 💜 available on the night!)? How 😏 ?! Submit 💻 your favourite <a href="https://twitter.com/hashtag/R?src=hash&amp;ref_src=twsrc%5Etfw">#R</a> tip to <a href="https://t.co/QI8F3FPAti">https://t.co/QI8F3FPAti</a> and present 🎬 it at our next event! Email also works! See here for more details <a href="https://t.co/M9nuKHWsVx">https://t.co/M9nuKHWsVx</a> <a href="https://t.co/vqSB2SJ9vT">pic.twitter.com/vqSB2SJ9vT</a></p>&mdash; R-Ladies Melbourne (@RLadiesMelb) <a href="https://twitter.com/RLadiesMelb/status/1047270616087162880?ref_src=twsrc%5Etfw">October 2, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
  
## Panel discussion
  
After the talks we also organised our first panel discussion with four R-Ladies from Melbourne at different stages of their career: Belinda Maher from Transport for Victoria, Nikki Rubinstein from PwC and Melissa Davis from Walter and Eliza Hall Institute.


<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Great <a href="https://twitter.com/RLadiesMelb?ref_src=twsrc%5Etfw">@RLadiesMelb</a> Sesh! So many awesome tips! <a href="https://t.co/J4Qiv3CcYZ">pic.twitter.com/J4Qiv3CcYZ</a></p>&mdash; Brigitte Lewis (@briglewis) <a href="https://twitter.com/briglewis/status/1052469031188353024?ref_src=twsrc%5Etfw">October 17, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>
  
Our panelist tackled the topics of workflows and public speaking. Melissa told us that her lab works with many different workflows depending on whether the research is aimed at exploratory analysis or method implementation. She explained how the package pkgdown is useful to showcase and share packages developed in her lab. She shared with the audience that in her early career as a researcher she actively volunteered to give talks at different seminars. This helped her research to reach a wider audience and ultimately meant that Melissa gained recognition in her chosen field. Belinda’s advice was to remember that when you are giving a public talk you need to make sure that the first thing the audience knows is why they should care about what you are telling them. You need to get their attention! Also Nikki’s go-to workflow is to set up a new R project for a new analysis and link it to a git repository, creating an organised folder structure for data, analysis files, reports and figures. The speakers also commented about the importance of always celebrating achievements! Nikki’s favourite celebration is going for frozen yoghurt! What’s yours?!


# First `#rstats` online lunch meetup
  
In mid-November, Soroor Hediyeh Zadeh suggested the great idea of starting an #rstats seminar series that we could hold as an alternative to or together with our monthly evening and longer meetups. Several members had already asked us to run morning or lunch events to suit other people’s and families’ needs. Well, there was the opportunity! Soroor suggested that these lunch seminars would be more focused on Statistics theory with, of course, R implementation. Initially we were worried that we couldn’t find a place with such short notice to host an event and that it might be hard for people to reach it when only a lunch break was available. That’s when we decided to try our first online Zoom meetup! That cut all costs and worries for us and allowed everyone to reach us in a click, sitting at their desk. While Soroor was presenting we muted the sound coming from every participant connected to the call to avoid background noise. Participants were still able to ask questions via the Zoom chat. It was our first online experiment and we did agree that it was a brilliant and successful one! It felt different and also we are still not sure how many people we are able to have in the same Zoom chat at once but it gave us the chance to quickly set up a meetup without needing to find a space and at no cost! Of course, the drawback is that you lose a bit the networking part of the meetup since it gets harder to network online, especially if you have a few people connected at the same time. I believe that alternating such events is key. We hope this will give a chance to a different audience to actively participate in the events and ask questions, and not just watch recorded events by themselves. 


<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr"><a href="https://twitter.com/hashtag/nonparametric?src=hash&amp;ref_src=twsrc%5Etfw">#nonparametric</a> vs <a href="https://twitter.com/hashtag/parametric?src=hash&amp;ref_src=twsrc%5Etfw">#parametric</a> models in <a href="https://twitter.com/hashtag/R?src=hash&amp;ref_src=twsrc%5Etfw">#R</a> by <a href="https://twitter.com/SoroorHediyeh?ref_src=twsrc%5Etfw">@SoroorHediyeh</a> through <a href="https://twitter.com/hashtag/zoom?src=hash&amp;ref_src=twsrc%5Etfw">#zoom</a> <a href="https://twitter.com/hashtag/online?src=hash&amp;ref_src=twsrc%5Etfw">#online</a> <a href="https://twitter.com/hashtag/rstat?src=hash&amp;ref_src=twsrc%5Etfw">#rstat</a> <a href="https://twitter.com/hashtag/RLadies?src=hash&amp;ref_src=twsrc%5Etfw">#RLadies</a> <a href="https://t.co/AisVGnvjez">pic.twitter.com/AisVGnvjez</a></p>&mdash; R-Ladies Melbourne (@RLadiesMelb) <a href="https://twitter.com/RLadiesMelb/status/1070122679238180865?ref_src=twsrc%5Etfw">December 5, 2018</a></blockquote>
  <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


# Get involved and share your ideas! - How to contact us
  
We are always looking for new speakers for seminars, workshops and new ideas! If you would like to get involved as organiser or in other ways, we would love to hear from you!
  
You can use any of the means below to contact us for any questions even if you need some advice or you have other questions about organising events in your city!
  
* `r icon::fa("twitter",colour = "#88398A")` https://twitter.com/RLadiesMelb
* `r icon::fa("github",colour = "#88398A")` https://github.com/R-LadiesMelbourne
* `r icon::fa("meetup",colour = "#88398A")` https://www.meetup.com/R-Ladies-Melbourne/
* `r icon::fa("facebook",colour = "#88398A")` https://www.facebook.com/RladiesMelbourne/
* Email: melbourne@rladies.org


# Testimonial from a past speaker: Belinda Maher

Belinda Maher was our speaker in [February 2018](https://www.youtube.com/watch?v=84G6ZNTEHCo&t=345s) and we also invited her for the panel discussion in October. Below you can find Belinda’s thoughts on how being a speaker at our event helped her gain confidence in herself and her skills!
  
*Presenting at R Ladies helped me to consider the audience's point of view: to think about exactly what was interesting and important about my work, and what needed further explanation for an audience who are not already familiar with the topic. I also gained presenting experience and had to think about how to change the content of my presentation to suit a less transport-focused audience than I am used to. While I notice that I sounded nervous at the start of my presentation, I would not hesitate to present my work again to a large group. The success of the presentation gave me confidence that I am quite capable of presenting and that it won’t be a disaster! Confidence that I will actually probably get good feedback if I present again. I discovered that it’s quite easy to talk about something which you’re passionate about. The networking opportunities which speaking at R Ladies has given me have also been a bonus. It’s nice also casually drop in to conversation when talking about my work: "oh, I did a presentation on that to 70 people at a user group, it’s on YouTube!"*


# Aknowledgements

* To produce this document I used some of the tips in [Pimp my Rmd](https://holtzy.github.io/Pimp-my-rmd/#use_dt_for_tables) by [Yan Holtz](https://github.com/holtzy)
* Thanks to my friend Steven for proof reading my first blog post! Thanks to Saskia, Sepideh and Marie for their helpful suggestions!
